---
title: "Developing Data Products - Week2 Assignment"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## <u><b> Objective </u></b> 
 - <u><b> The objectives of this document are: </u></b> 
    * Create an interactive map using Leaflet package
    * Create an R Markdown document embedding the interactive map
    * Host the R Markdown document on GitHub Pages
    
## <u><b> "Quakes" Dataset - Exploration </u></b> 

#### <u><b> Loading "Quakes" Dataset </u></b> 
 - Lets load the R built in "quakes" dataset
 
```{r load_dataset}
 library(leaflet)
 library(dplyr)
 data("quakes")
```

#### <u><b> "Quakes" Dataset - Inpection </u></b> 
 - Lets look at first few rows of "quakes" dataset via head() command
 
```{r view_dataset}
 head(quakes)
```

 - Lets look at unique values of mag column which signifies magnitude of the earth quake
 
```{r view_mag}
 unique(quakes$mag)
```

 - Magnitudes are all in ranges (4.0 ~ 4.9), (5.0 ~ 5.9), (6.0 ~ 6.9)

#### <u><b> "Quakes" Dataset - Enrichment </u></b> 
 - Lets assign some colors (add a new column) as per the below guidelines. This would help later in mapping this
    * (4.0 ~ 4.9) -> Blue
    * (5.0 ~ 5.9) -> Orange
    * (6.0 ~ 6.9) -> Red
 
```{r add_color}
   quakes <- quakes %>% mutate (
     color = case_when(
         mag >= 4.0 & mag <= 4.9 ~ "blue",
         mag >= 5.0 & mag <= 5.9 ~ "orange",
         mag >= 6.0 & mag <= 6.9 ~ "red",
     )
   )
```

#### <u><b> "Quakes" Dataset - Subset </u></b> 
 - Lets group the rows by color, sort the group in descending order by their magnitudes and pick top 15 rows
 - We are trying to restrict analysis on the top 15 events among each group
 
```{r group_dataset}
 quakes <- quakes %>%
    group_by(color) %>%
    arrange(desc(mag)) %>%
    do(head(., n = 15))
```

 - Lets look at first few rows of "quakes" dataset again to make sure we have the new color column
 
```{r view_quakes_dataset}
 quakes
```

## <u><b> "Quakes" Dataset - Interactive Maps </u></b> 
 - Lets plot the derived quakes dataset into an interative map via leaflet package
 
```{r map_dataset}
quakes %>% 
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(color = quakes$color, popup = paste("Magnitude = ", quakes$mag), radius = ~ mag ^ 2,
    stroke = FALSE, fillOpacity = 0.5) %>%
  addLegend(labels = c("Mag = (4.0 ~ 4.9)", "Mag = (5.0 ~ 5.9)", "Mag = (6.0 ~ 6.9)"), colors = c("blue", "red", "green"))
```

 - As the legend says:
    - Red circles signifies earthquakes reaching magnitude of (6.0 ~ 6.9)
    - Orange circles signifies earthquakes reaching magnitude of (5.0 ~ 5.9)
    - Blue circles signifies earthquakes reaching magnitude of (4.0 ~ 4.9)
    
 - The radius of the circle markers increases with the increasing value of magnitude
    - Hence Red circles are larger compared to Blue ones
    
 - Clicking on any cirlce would yield a pop up to display it's magnitude value